One page description

### 2016-09-30 publications/Journal/6-allocation.tex
Address first referee's comments (p1)
Signed-off-by: Ben Cardoen <bencardoen@gmail.com>

Although the synchronization protocol is one of the defining factors in simulation performance, model allocation has a significant impact on which protocol is ideal.
Depending on the model structure, and how it is mapped to the different cores, it might not even make sense to parallelize at all.
Indeed, if the model is distributed such that frequent communication is necessary between cores, parallelism is naturally reduced.
This brings us to the topic of model allocation, as also implemented in \textit{PythonPDEVS}~\cite{PythonPDEVS2}.

@@

@@

After simulation the traces can be visualized for both breadth-first and depth-first allocation.
Using a breadth-first allocation scheme, as shown in Figure~\ref{fig:pholdtree_visualize_parBFS}, we notice that many events get exchanged between cores.
This is caused by the high number of inter-core connections and the high number of events exchanged over these connections.
The number of connections between nodes at the same simulation core is also rather low.
Using a depth-first allocation scheme, however, as shown in Figure~\ref{fig:pholdtree_visualize_parDFS}, minimizes inter-core connections while maximizing intra-core connections.

@@

The most prominent aspect of these results is the low performance for conservative depth-first allocation for two cores.
This is mostly caused by the difference between a sequential simulation and a parallel simulation: suddenly we need to take into account synchronization and passing around of lookahead values.
And since the number of cores is low, the overhead dominates any further speedup.
Optimistic is less sensitive to the number of models per core as it does not need to poll each model for a lookahead, this explains the lower runtime penalty observed for optimistic.
%TODO add some more explanation here maybe?
%TODO why is this not the case for optimistic?

%TODO this paragraph is strange: is this really the case? 
%The effect is due to the topology of the kernels, in BFS the risk on reverts is higher. Optimistic is less influenced but will get a lower speedup in the good case.
Interestingly, we see that optimistic synchronization is less influenced by the allocation than conservative synchronization.
This is likely caused due to the lower number of connections to take into account in conservative synchronization.
Whereas conservative synchronization needs to take into account even scarcely used connections, optimistic synchronization does not.
The same is true in the opposite direction, though, where optimistic synchronization is slower when a good allocation is chosen.
Conservative synchronization will then be able to make better estimates, whereas optimistic synchronization does not make estimations.

>>>>

Although the synchronization protocol is one of the defining factors in simulation performance, model allocation has a significant impact on which protocol is ideal.
Depending on the model structure, and how it is mapped to the different kernels, it might not even make sense to parallelize at all.
Indeed, if the model is distributed such that frequent communication is necessary between kernels, parallelism is naturally reduced.
This brings us to the topic of model allocation, as also implemented in \textit{PythonPDEVS}~\cite{PythonPDEVS2}.

@@

Both allocators will try to divide models evenly over kernels. The effects of varying the number of models per kernel are already evaluated in the previous section on scaling. Here we want to highlight the overhead of communication and inter-kernel dependence.

@@

After simulation the traces can be visualized for both breadth-first and depth-first allocation.
Using a breadth-first allocation scheme, as shown in Figure~\ref{fig:pholdtree_visualize_parBFS}, we notice that many events get exchanged between kernels.
This is caused by the high number of inter-kernel connections and the high number of events exchanged over these connections.
The number of connections between nodes at the same simulation kernel is also rather low.
Using a depth-first allocation scheme, however, as shown in Figure~\ref{fig:pholdtree_visualize_parDFS}, minimizes inter-kernel connections while maximizing intra-kernel connections.

@@

The most prominent aspect of these results is the low performance for conservative depth-first allocation for two kernels.
This is mostly caused by the difference between a sequential simulation and a parallel simulation: suddenly we need to take into account synchronization and passing around of lookahead values.
And since the number of kernels is low, the overhead dominates any further speedup.
Optimistic is less sensitive to the number of models per kernel as it does not need to poll each model for a lookahead, this explains the lower runtime penalty observed for optimistic.

%The effect is due to the topology of the kernels, in BFS the risk on reverts is higher. Optimistic is less influenced but will get a lower speedup in the good case.
Interestingly, we see that optimistic synchronization is less influenced by the allocation than conservative synchronization.
This is likely caused due to the lower number of connections to take into account in conservative synchronization.
Whereas conservative synchronization needs to take into account even scarcely used connections, optimistic synchronization does not.
The same is true in the opposite direction, though, where optimistic synchronization is slower when a good allocation is chosen.
Conservative synchronization will then be able to make better estimates, whereas optimistic synchronization does not make estimations.

### 2016-09-30 publications/Journal/memory.tex

\subsection{Memory Usage}
Apart from simulation execution time, memory usage during simulation is also of great importance.
While execution time only becomes a problem if it takes way too long, coming short only a bit of memory can make simulation unfeasible.
We therefore also investigate memory usage of different synchronization protocols, and again compare to \textit{adevs}.

@@

\subsubsection{Remarks}
Both \textit{dxex} and \textit{adevs} use \textit{tcmalloc} as memory allocator, allowing for thread-local allocation.
Additionally, dxex uses memory pools to further reduce the frequency of expensive system calls (\textit{e.g.}, malloc and free).
\textit{Tcmalloc} only gradually releases memory back to the OS, whereas our pools will not do so at all.
Due to our motivation for memory usage analysis, we will only measure peak allocation in maximum resident set size as reported by the OS.

>>>>

\subsection{Memory Usage}
Apart from simulation execution time, memory usage during simulation is also of great importance.
While execution time only becomes a problem if it takes way too long, coming short only a bit of memory can make simulation infeasible.
We therefore also investigate memory usage of different synchronization protocols, and again compare to \textit{adevs}.

@@

\subsubsection{Remarks}
Both \textit{dxex} and \textit{adevs} use \textit{tcmalloc} as memory allocator, allowing for thread-local allocation.
Additionally, dxex uses memory pools to further reduce the frequency of expensive system calls (\textit{e.g.}, malloc and free).
\textit{tcmalloc} only gradually releases memory back to the OS, whereas our pools will not do so at all.
Due to our motivation for memory usage analysis, we will only measure peak allocation in maximum resident set size as reported by the OS.

### 2016-09-30 publications/Journal/models.tex

Second, the \textit{PHOLD} model~\cite{PHOLD}, creates $n$ atomic models, where each model has exactly $n-1$ output ports.
Each atomic model is directly connected to every other atomic model.
After a random delay, an atomic model sends out an event to a randomly selected output port.
Output port selection happens in two phases: first it is decided whether the event should be sent within the same node, or outside of the node.
Afterwards, a uniform selection is made between the remaining models.
The model takes two parameter: the percentage of remote events, which determines the fraction of messages routed to other nodes, and the percentage of priority events. Priority events are events generated in a very short time after the previous event.
This benchmark shows how the simulation kernel behaves in the presence of many local or remote events, in combination with a varying percentage of high-priority event.s
An example for four models, split over two nodes, is shown in Figure~\ref{fig:PHOLD_model}.

Third, the \textit{Interconnect} model, a merge of PHOLD~\cite{PHOLD} and the \textit{HI} model of DEVStone~\cite{DEVStone}, creates $n$ atomic models, where each model has exactly one output port.
Similar to PHOLD, all models are connected to one another, but all through the same port: every model receives each generated event.
The model takes one parameter: the number of models.
This benchmark shows the complexity of event creation, event routing, and simultaneous event handling.
An example for four models is shown in Figure~\ref{fig:interconnect_model}.

>>>>

Second, the \textit{Interconnect} model, a merge of PHOLD~\cite{PHOLD} and the \textit{HI} model of DEVStone~\cite{DEVStone}, creates $n$ atomic models, where each model has exactly one output port.
Similar to PHOLD, all models are connected to one another, but all through the same port: every model receives each generated event.
The model takes one parameter: the number of models.
This benchmark shows the complexity of event creation, event routing, and simultaneous event handling.
An example for three models is shown in Figure~\ref{fig:interconnect_model}.

Third, the \textit{PHOLD} model~\cite{PHOLD}, creates $n$ atomic models, where each model has exactly $n-1$ output ports.
Each atomic model is directly connected to every other atomic model.
After a random delay, an atomic model sends out an event to a randomly selected output port.
Output port selection happens in two phases: first it is decided whether the event should be sent within the same node, or outside of the node.
Afterwards, a uniform selection is made between the remaining models.
The model takes two parameter: the percentage of remote events, which determines the fraction of messages routed to other nodes, and the percentage of priority events. Priority events are events generated in a very short time after the previous event.
This benchmark shows how the simulation kernel behaves in the presence of many local or remote events, in combination with a varying percentage of high-priority event.s
An example for four models, split over two nodes, is shown in Figure~\ref{fig:PHOLD_model}.
